### 2016-04-14 figure_2.Rmd

---
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = FALSE, message = FALSE)
```

```{r}

# Directories.
out.dir <- "OUT"

nsT.theo.MSDi.s = c()
exp.MSDi.s = c()

# Read input.
input.fname <- file.path("input_MainMultipleReport.csv")
input <- read.csv(input.fname)

# Start a loop to read data for each family.
for (a in (1:nrow(input))) { 
  family <- as.character(input$family)[a]
  mut.model = input$mut.model[a]
  R0 = input$R0[a]
  K.analysis = input$K.analysis[a]

  #Read experimental data.
  analysis.fname.id <- paste(family, "_mut.model_", mut.model, "_naturalSelection_", "TRUE", "_R0_", R0, "_core_", "FALSE", "_K.analysis_", K.analysis, sep = "")

  m.cF.exp.smooth.dr.squarei = read.csv(file.path(out.dir, paste(analysis.fname.id, "_out_m.exp.smooth.dr.squarei.csv", sep = "")))

  cF.exp.MSDi.s = colMeans(m.cF.exp.smooth.dr.squarei, na.rm = T )

  #Read theoretical data#
  m.nsT.cF.theo.smooth.dr.squarei = read.csv(file.path(out.dir, paste(analysis.fname.id, "_out_m.theo.smooth.dr.squarei.csv", sep = "")))

  nsT.cF.theo.MSDi.s = colMeans(m.nsT.cF.theo.smooth.dr.squarei, na.rm = T)

  ## Fit theoretical data with experimental data.

  fit.nsT.cF.s = lm(cF.exp.MSDi.s ~ nsT.cF.theo.MSDi.s)
  fit.nsT.cF.MSDi.s = fitted.values(fit.nsT.cF.s)

  # Concatenate information.

  nsT.theo.MSDi.s = c(nsT.theo.MSDi.s, fit.nsT.cF.MSDi.s)

  cF.exp.MSDi.s = cF.exp.MSDi.s[!is.na(cF.exp.MSDi.s)]
  exp.MSDi.s = c(exp.MSDi.s, cF.exp.MSDi.s)
}

# Caluclate measures.

R2.nsT.cF.s = cor(nsT.theo.MSDi.s, exp.MSDi.s) ^ 2

MSE.nsT.cF.s = mean((nsT.theo.MSDi.s - exp.MSDi.s) ^ 2)

```

```{r}
plot(nsT.theo.MSDi.s, type = "l", xlab = "site", ylab = "Theoretical MSDi NS = T", ylim = c(0, 0.14))
```

```{r}
plot(exp.MSDi.s, type = "l", col = "red", xlab = "site",  ylab = "Experimental MSDi ", ylim = c(0, 0.14))
```

```{r}
plot(nsT.theo.MSDi.s, type = "l", xlab = "site", ylab = "Merge", ylim = c(0, 0.14))
points(exp.MSDi.s, col = "red", type = "l")
legend("topleft", paste("R^2", round(R2.nsT.cF.s, digits = 2), " ", "MSE", signif(MSE.nsT.cF.s, digits = 4)), bty = "n", text.col = "red")
```