### 2017-04-23 analysis-structure-core.Rmd

---
title: "Structural Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

# Introduction
We study the evolutionary divergence of proteins using the LF-ENM and comparing simulated mutants with experimental data. 
For the core we considered sites with no gaps in the whole multiple alignment.

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

# load functions
source(file.path(data.dir, "multiplot.R"))
source(file.path(data.dir, "my-functions.R"))
```

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "medium", "selection")
```

```{r set-run-variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))
sum.inv.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_sum.inv.da.CM.ca.csv", sep = ""))

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
sim.data.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
sim.data.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_out_df.data.csv", sep = ""))
```

```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set-experimental-protein-names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r read-the-core, comment = ""}
# indexes.core
indexes.core = as.numeric(read.csv(file.path("DATA", paste("indexes.core", "_", family, ".csv", sep = "")))[1, -1])
```

```{r set-site-info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)
m.sum.inv.da.CM.CA = read.csv(sum.inv.da.CM.CA.fname)

# add info to site.info
d = ddply(site.info, "site", .fun = function(x) data.frame("da" = min(x[, c(-1, -2)])))

# define shells of active site neighborhood
d$shell = cut(d$da, breaks = c(-.1, 2.5, 7.5, 12.5, 17.5, 22.5, 27.5, 32.5, 37.5, 42.5, 47.5, 52.5, max(d$da)))
levels(d$shell) = seq(length(levels(d$shell))) - 1 # rename shell levels
site.info = merge(site.info, d, by = "site")

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

# add pdb.info to site.info
site.info = cbind(site.info[indexes.core, ], pdb.info[indexes.core, ])
```

```{r read-data}
# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)[, indexes.core]
rownames(dri2.exp) = protein.exp

## mut
dri2.mut = read.csv(dri2.mut.fname, header = T)[, indexes.core]
rownames(dri2.mut) = paste("mut", rownames(dri2.mut), sep = "")

## medium
dri2.medium = read.csv(dri2.medium.fname, header = T)[, indexes.core]
rownames(dri2.medium) = paste("medium", rownames(dri2.medium), sep = "")
```

# Structural divergence along sites

Here we study how structural divergence varies from site to site. We compare simulated data with empirical data.

```{r prepare-site-data}
# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## mut
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.mut = cbind(site.info, tdri2.mut)
siteComparison.mut = melt(tdri2.mut, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## medium
tdri2.medium = as.data.frame(t(as.matrix(dri2.medium)))
tdri2.medium = cbind(site.info, tdri2.medium)
siteComparison.medium = melt(tdri2.medium, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "medium", siteComparison.medium))

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "n.SD" = dri2/mean(dri2, na.rm = T),
          "z.SD" = vscale(dri2),
          "RSD" = sqrt(dri2), 
          "n.RSD" = RSD/mean(RSD, na.rm = T),
          "z.RSD" = vscale(RSD))
```

### Mean z.RSD vs site

To measure structural divergence we use z.RSD profiles, which are the best for most families.

The following plot compares the mean z.RSD profiles, including standard errors (with .999 confidence level for the mean)
*Warning: the effective number of data should be smaller than the actual data, especially for the experimental dataset because of several factors. We need to deal with this for quantitative analysis. Probably we'll need to use bootstrapping to calculate standard errors non-parametrically*

```{r mean-se-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))

## dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})

## find the active site and its contacts
active.sites = d$site[d$shell == 0]
contacts.active.sites = d$site[d$shell == 1]

## prepare plots with mean profiles and deviations
p = ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_point() +
    geom_errorbar(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(<z.RSD>)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

## plot separate and comparative plots
p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "bottom")

p.facet %+% subset(d, dataset == "exp") 
p.facet %+% subset(d, dataset == "no-selection")
p.facet %+% subset(d, dataset == "selection") 
p.facet

## plot superimposed plots
p + theme(legend.position = "bottom")
```

### Correlation between site profiles

```{r comparison-mean-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}

# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

# eliminate duplicates
active.sites = active.sites[!duplicated(active.sites)]
contacts.active.sites = contacts.active.sites[!duplicated(active.sites)]

# new.indexes
new.active.index = c()
for (i in (1:(length(active.sites)))){
  new.active.index[i] = which(indexes.core == active.sites[i])
}

new.contacts.active.index = c()
for (i in (1:(length(contacts.active.sites)))){
  new.contacts.active.index[i] = which(indexes.core == contacts.active.sites[i])
}

active.sites = new.active.index
contacts.active.sites = new.contacts.active.index

# calculate cors
r.exp.mut = cor(d.exp$mean.z.RSD, d.mut$mean.z.RSD)
r.exp.medium = cor(d.exp$mean.z.RSD, d.medium$mean.z.RSD)
r.mut.medium = cor(d.mut$mean.z.RSD, d.medium$mean.z.RSD)

cor.df = data.frame("comparison" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))

knitr::kable(cor.df, digits = 2, caption = "Correlation between <Z-scores> profiles")
```

### Comparison between profiles

```{r comparisons-between-profiles}

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = d.exp$mean.z.RSD, 
     y = d.mut$mean.z.RSD, 
     xlab = "<z.RSD> experimental", 
     ylab = "<z.RSD> no-selection", 
     xlim = c(min(c(d.exp$mean.z.RSD, d.mut$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.mut$mean.z.RSD))),
     ylim = c(min(c(d.exp$mean.z.RSD, d.mut$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.mut$mean.z.RSD))))
points(x = as.numeric(d.exp$mean.z.RSD)[active.sites], y = as.numeric(d.mut$mean.z.RSD)[active.sites], col = "yellow")
points(x = as.numeric(d.exp$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.mut$mean.z.RSD)[contacts.active.sites], col = "green")
abline(0, 1)

plot(x = d.exp$mean.z.RSD, 
     y = d.medium$mean.z.RSD, 
     xlab = "<z.RSD> experimental", 
     ylab = "<z.RSD> selection", 
     xlim = c(min(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD))),
     ylim = c(min(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD))))
points(x = as.numeric(d.exp$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.medium$mean.z.RSD)[contacts.active.sites], col = "green")
points(x = as.numeric(d.exp$mean.z.RSD)[active.sites], y = as.numeric(d.medium$mean.z.RSD)[active.sites], col = "yellow")
abline(0, 1)

plot(x = d.mut$mean.z.RSD, 
     y = d.medium$mean.z.RSD, 
     xlab = "<z.RSD> no-selection", 
     ylab = "<z.RSD> selection", 
     xlim = c(min(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD))),
     ylim = c(min(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD))))
points(x = as.numeric(d.mut$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.medium$mean.z.RSD)[contacts.active.sites], col = "green")
points(x = as.numeric(d.mut$mean.z.RSD)[active.sites], y = as.numeric(d.medium$mean.z.RSD)[active.sites], col = "yellow")
abline(0, 1)
```
### Distance to the active site

```{r comparison-min.da}

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = (d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], x = m.min.da.CM.CA$min.da.CA[indexes.core][-active.sites], xlab = "minimal distance to active site (between CAs)", ylab = "<z.RSD> experimental - <z.RSD> no-selection")
plot(y = (d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], x = m.min.da.CM.CA$min.da.CM[indexes.core][-active.sites], xlab = "minimal distance to active site (between CMs)", ylab = "<z.RSD> experimental - <z.RSD> no-selection")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], m.min.da.CM.CA$min.da.CA[indexes.core][-active.sites])
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], m.min.da.CM.CA$min.da.CM[indexes.core][-active.sites])

cor.df = data.frame("comparison" = c("(<z.RSD> experimental - <z.RSD> no-selection) vs min.da.CA", "(<z.RSD> experimental - <z.RSD> no-selection) vs min.da.CM"), "R" = c(r.min.da.CA, r.min.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the minimal distance to the active site")

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = (d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], x = m.min.da.CM.CA$min.da.CA[indexes.core][-active.sites], xlab = "minimal distance to active site (between CAs)", ylab = "<z.RSD> experimental - <z.RSD> selection")
plot(y = (d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], x = m.min.da.CM.CA$min.da.CM[indexes.core][-active.sites], xlab = "minimal distance to active site (between CMs)", ylab = "<z.RSD> experimental - <z.RSD> selection")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], m.min.da.CM.CA$min.da.CA[indexes.core][-active.sites])
r.min.da.CM = cor((d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], m.min.da.CM.CA$min.da.CM[indexes.core][-active.sites])

cor.df = data.frame("comparison" = c("(<z.RSD> experimental - <z.RSD> selection) vs min.da.CA", "(<z.RSD> experimental - <z.RSD> selection) vs min.da.CM"), "R" = c(r.min.da.CA, r.min.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the minimal distance to the active site")
```

```{r comparison-sum.inv.da}

# comparison of means.z.scores with the sum.inv distance to the functional sites
plot(y = (d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], x = m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core][-active.sites], xlab = "sum.inv distance to active site (between CAs)", ylab = "<z.RSD> experimental - <z.RSD> no-selection")
plot(y = (d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], x = m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core][-active.sites], xlab = "sum.inv distance to active site (between CMs)", ylab = "<z.RSD> experimental - <z.RSD> no-selection")

## calculate correlations between profiles
r.sum.inv.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core][-active.sites])
r.sum.inv.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD)[-active.sites], m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core][-active.sites])

cor.df = data.frame("comparison" = c("(<z.RSD> experimental - <z.RSD> no-selection) vs sum.inv.da.CA", "(<z.RSD> experimental - <z.RSD> no-selection) vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA, r.sum.inv.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the sum.inv distance to the active site")

# comparison of means.z.scores with the sum.inv distance to the functional sites
plot(y = (d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], x = m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core][-active.sites], xlab = "sum.inv distance to active site (between CAs)", ylab = "<z.RSD> experimental - <z.RSD> selection")
plot(y = (d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], x = m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core][-active.sites], xlab = "sum.inv distance to active site (between CMs)", ylab = "<z.RSD> experimental - <z.RSD> selection")

## calculate correlations between profiles
r.sum.inv.da.CA = cor((d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core][-active.sites])
r.sum.inv.da.CM = cor((d.exp$mean.z.RSD - d.medium$mean.z.RSD)[-active.sites], m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core][-active.sites])

cor.df = data.frame("comparison" = c("(<z.RSD> experimental - <z.RSD> selection) vs sum.inv.da.CA", "(<z.RSD> experimental - <z.RSD> selection) vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA, r.sum.inv.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the sum.inv distance to the active site")
```

```{r correlation-da-CN} 
## CN core
CN = CN[indexes.core]

## calculate correlations between profiles
r.min.da.CA.CN = cor(CN[-active.sites], m.min.da.CM.CA$min.da.CA[indexes.core][-active.sites])
r.min.da.CM.CN = cor(CN[-active.sites], m.min.da.CM.CA$min.da.CM[indexes.core][-active.sites])

cor.df = data.frame("comparison" = c("CN vs min.da.CA", "CN vs min.da.CM"), "R" = c(r.min.da.CA.CN, r.min.da.CM.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CN and the minimal distance to the active site")

## calculate correlations between profiles
r.sum.inv.da.CA.CN = cor(CN[-active.sites], m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core][-active.sites])
r.sum.inv.da.CM.CN = cor(CN[-active.sites], m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core][-active.sites])

cor.df = data.frame("comparison" = c("CN vs sum.inv.da.CA", "CN vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA.CN, r.sum.inv.da.CM.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CN and the sum.inv distance to the active site")
```

### Sequence comparison

In this section we compare theoretical sequence divergence profiles with sequence scores obtained from ConsurfDB. To obtain theoretical profiles, for each selection regimen, we calculated the number of times we had mutated each site and divided this number by the number of times we had tried a mutation at the site.

```{r sequence-profiles-comparison}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE[indexes.core]

# read simulated data
## no sel
sim.data.mut = read.csv(sim.data.mut.fname)
mut.site.mut = sim.data.mut$site
accept.mut = sim.data.mut$accept

## medium sel
sim.data.medium.sel = read.csv(sim.data.medium.sel.fname)
mut.site.medium.sel = sim.data.medium.sel$site
accept.medium.sel = sim.data.medium.sel$accept

# calculate p.accept of simulated data
p.accept.mut = c()
p.accept.medium.sel = c()

if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.mut = mut.site.mut - n.sites
  mut.site.medium.sel = mut.site.medium.sel - n.sites
}

# core n.sites

for (i in (1:n.sites)) {
  accept.mut.i = accept.mut[which(mut.site.mut == i)]
  p.accept.mut.i = mean(accept.mut.i)
  p.accept.mut[i] = p.accept.mut.i

  accept.medium.sel.i = accept.medium.sel[which(mut.site.medium.sel == i)]
  p.accept.medium.sel.i = mean(accept.medium.sel.i)
  p.accept.medium.sel[i] = p.accept.medium.sel.i
}
p.accept.mut = p.accept.mut[indexes.core]
p.accept.medium.sel = p.accept.medium.sel[indexes.core]

# compare exp vs no sel profiles
## plot
par(pty = "s")

plot(x = p.accept.mut, y = CN, xlab = "log(paccept no-selection)")
plot(x = p.accept.mut, y = score, xlab = "log(paccept no-selection)")
plot(x = score, y = CN)

## calculate corelations
r.p.accept.mut.CN = cor(log(p.accept.mut), CN)
r.p.accept.mut.score = cor(log(p.accept.mut), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("log(paccept no-selection) vs CN", "log(paccept no-selection) vs score", "score vs CN"),
                    "R" = c(r.p.accept.mut.CN , r.p.accept.mut.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

# compare exp vs medium sel profiles
## plot
par(pty = "s")

plot(x = log(p.accept.medium.sel), y = CN, xlab = "log(paccept selection)")
plot(x = log(p.accept.medium.sel), y = score, xlab = "log(paccept selection)")
plot(x = score, y = CN)

## calculate corelations
r.p.accept.medium.sel.CN = cor(log(p.accept.medium.sel), CN)
r.p.accept.medium.sel.score = cor(log(p.accept.medium.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("log(paccept selection) vs CN", "log(paccept selection) vs score", "score vs CN"),
                    "R" = c(r.p.accept.medium.sel.CN , r.p.accept.medium.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

```