### 2017-04-27 comparison_CA_CM.Rmd
CA vs CM

---
title: "Structural Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load-packages}
# load functions
source(file.path(data.dir, "my-functions.R"))

# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)
```

### Structure - cartesian coordinates

```{r set-file-names}
data.dir <- paste("OUT/out_subset_CA_ANM", sep = "")
R0 = R0.CA

# set input filenames
## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
CA.sim.data.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
CA.sim.data.medium.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_out_df.data.csv", sep = ""))

## active sites
functional.sites.fname = file.path("DATA", paste(p.ref, "_functionalSites.csv", sep = ""))

```

```{r read-functional-sites}
# which are the functional sites?
functional.sites = read.csv(functional.sites.fname, sep = ";")
active.site.index = functional.sites$index
```

```{r read-data}
# read data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)

n.sites = length(dri2.exp)

rownames(dri2.exp) = paste("mut", rownames(dri2.exp), sep = "")
rownames(dri2.mut) = paste("mut", rownames(dri2.mut), sep = "")
rownames(dri2.medium) = paste("mut", rownames(dri2.medium), sep = "")

tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.medium = as.data.frame(t(as.matrix(dri2.medium)))

site.info = seq(1:length(dri2.mut))

tdri2.exp = cbind(site.info, tdri2.exp)
tdri2.mut = cbind(site.info, tdri2.mut)
tdri2.medium = cbind(site.info, tdri2.medium)

siteComparison.exp = melt(tdri2.exp, id.vars = c("site.info"), variable.name = "protein", value.name = "dri2")
siteComparison.mut = melt(tdri2.mut, id.vars = c("site.info"), variable.name = "protein", value.name = "dri2")
siteComparison.medium = melt(tdri2.medium, id.vars = c("site.info"), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "medium", siteComparison.medium))

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "n.SD" = dri2/mean(dri2, na.rm = T),
          "z.SD" = vscale(dri2),
          "RSD" = sqrt(dri2),
          "n.RSD" = RSD/mean(RSD, na.rm = T),
          "z.RSD" = vscale(RSD))

dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

## dat with means and standard errors
d.CA = ddply(dat,c("dataset", "site.info"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})
```

```{r set-file-names-CM}
data.dir <- paste("OUT/out_subset_CM_ANM", sep = "")
R0 = R0.CM

# set input filenames
## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
CM.sim.data.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
CM.sim.data.medium.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_out_df.data.csv", sep = ""))
```

```{r read-data-CM}
# read data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)

rownames(dri2.exp) = paste("mut", rownames(dri2.exp), sep = "")
rownames(dri2.mut) = paste("mut", rownames(dri2.mut), sep = "")
rownames(dri2.medium) = paste("mut", rownames(dri2.medium), sep = "")

tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.medium = as.data.frame(t(as.matrix(dri2.medium)))

site.info = seq(1:length(dri2.mut))

tdri2.exp = cbind(site.info, tdri2.exp)
tdri2.mut = cbind(site.info, tdri2.mut)
tdri2.medium = cbind(site.info, tdri2.medium)

siteComparison.exp = melt(tdri2.exp, id.vars = c("site.info"), variable.name = "protein", value.name = "dri2")
siteComparison.mut = melt(tdri2.mut, id.vars = c("site.info"), variable.name = "protein", value.name = "dri2")
siteComparison.medium = melt(tdri2.medium, id.vars = c("site.info"), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "medium", siteComparison.medium))

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "n.SD" = dri2/mean(dri2, na.rm = T),
          "z.SD" = vscale(dri2),
          "RSD" = sqrt(dri2),
          "n.RSD" = RSD/mean(RSD, na.rm = T),
          "z.RSD" = vscale(RSD))

dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

## dat with means and standard errors
d.CM = ddply(dat,c("dataset", "site.info"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})
```

```{r comparison-mean-z.RSD-vs-site}
# CA
d = d.CA

# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

# calculate cors
r.exp.mut = cor(d.exp$mean.z.RSD, d.mut$mean.z.RSD)
r.exp.medium = cor(d.exp$mean.z.RSD, d.medium$mean.z.RSD)
r.mut.medium = cor(d.mut$mean.z.RSD, d.medium$mean.z.RSD)

# save data
CA.mean.exp = d.exp$mean.z.RSD
CA.mean.mut = d.mut$mean.z.RSD
CA.mean.medium = d.medium$mean.z.RSD

CA.se.exp = d.exp$se.z.RSD
CA.se.mut = d.mut$se.z.RSD
CA.se.medium = d.medium$se.z.RSD

cor.df = data.frame("comparison CA" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))

knitr::kable(cor.df, digits = 2, caption = "Correlation between <Z-scores> profiles")

# CM
d = d.CM

# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

# calculate cors
r.exp.mut = cor(d.exp$mean.z.RSD, d.mut$mean.z.RSD)
r.exp.medium = cor(d.exp$mean.z.RSD, d.medium$mean.z.RSD)
r.mut.medium = cor(d.mut$mean.z.RSD, d.medium$mean.z.RSD)

# save data
CM.mean.exp = d.exp$mean.z.RSD
CM.mean.mut = d.mut$mean.z.RSD
CM.mean.medium = d.medium$mean.z.RSD

CM.se.exp = d.exp$se.z.RSD
CM.se.mut = d.mut$se.z.RSD
CM.se.medium = d.medium$se.z.RSD

cor.df = data.frame("comparison CM" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))

knitr::kable(cor.df, digits = 2, caption = "Correlation between <Z-scores> profiles")
```

```{r plots, fig.cap= "yellow = active sites"}
# plot comparative profiles
par(pty = "s")

# exp
plot(x = CA.mean.exp, y = CM.mean.exp, xlab = "CA <z.RSD> experimental", ylab = "CM <z.RSD> experimental")
abline(0, 1)
points(x = CA.mean.exp[active.site.index], y = CM.mean.exp[active.site.index], col = "yellow")

# mut
plot(x = CA.mean.mut, y = CM.mean.mut, xlab = "CA <z.RSD> no-selection", ylab = "CM <z.RSD> no-selection")
abline(0, 1)
points(x = CA.mean.mut[active.site.index], y = CM.mean.mut[active.site.index], col = "yellow")

# medium
plot(x = CA.mean.medium, y = CM.mean.medium, xlab = "CA <z.RSD> selection", ylab = "CM <z.RSD> selection")
abline(0, 1)
points(x = CA.mean.medium[active.site.index], y = CM.mean.medium[active.site.index], col = "yellow")
```

```{r correlations-CA-CM}
# calculate cors
r.exp = cor(CA.mean.exp, CM.mean.exp)
r.mut = cor(CA.mean.mut, CM.mean.mut)
r.medium = cor(CA.mean.medium, CM.mean.medium)

cor.df = data.frame("comparison CM" = c("exp CA vs CM", "no-selection CA vs CM", "selection CA vs CM"),
                    "R" = c(r.exp, r.mut, r.medium))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CA and CM <Z-scores> profiles")
```

```{r sequence-profiles-comparison}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

# read simulated data
## no sel
CA.sim.data.mut = read.csv(CA.sim.data.mut.fname)
mut.site.mut = CA.sim.data.mut$site
accept.mut = CA.sim.data.mut$accept

## medium sel
CA.sim.data.medium.sel = read.csv(CA.sim.data.medium.sel.fname)
mut.site.medium.sel = CA.sim.data.medium.sel$site
accept.medium.sel = CA.sim.data.medium.sel$accept

# calculate p.accept of simulated data
p.accept.mut = c()
p.accept.medium.sel = c()

# calculate CN
CN.CA = CA.sim.data.mut$CN
CN = c()

for (i in (1:n.sites)) {
  CN[i] = CN.CA[which(mut.site.mut == i)][1]
}

for (i in (1:n.sites)) {
  accept.mut.i = accept.mut[which(mut.site.mut == i)]
  p.accept.mut.i = mean(accept.mut.i)
  p.accept.mut[i] = p.accept.mut.i

  accept.medium.sel.i = accept.medium.sel[which(mut.site.medium.sel == i)]
  p.accept.medium.sel.i = mean(accept.medium.sel.i)
  p.accept.medium.sel[i] = p.accept.medium.sel.i
}

## calculate corelations
r.p.accept.mut.CN = cor(p.accept.mut, CN)
r.p.accept.mut.score = cor(p.accept.mut, score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison CA" = c("log(paccept no-selection) vs CN", "log(paccept no-selection) vs score", "score vs CN"),
                    "R" = c(r.p.accept.mut.CN , r.p.accept.mut.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

print("<paccept no-selection>")
print(mean(p.accept.mut))

## calculate corelations
r.p.accept.medium.sel.CN = cor(log(p.accept.medium.sel), CN)
r.p.accept.medium.sel.score = cor(log(p.accept.medium.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison CA" = c("log(paccept selection) vs CN", "log(paccept selection) vs score", "score vs CN"),
                    "R" = c(r.p.accept.medium.sel.CN , r.p.accept.medium.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

print("<paccept selection>")
print(mean(p.accept.medium.sel))
```

```{r sequence-profiles-comparison-CM}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

# read simulated data
## no sel
CM.sim.data.mut = read.csv(CM.sim.data.mut.fname)
mut.site.mut = CM.sim.data.mut$site
accept.mut = CM.sim.data.mut$accept

## medium sel
CM.sim.data.medium.sel = read.csv(CM.sim.data.medium.sel.fname)
mut.site.medium.sel = CM.sim.data.medium.sel$site
accept.medium.sel = CM.sim.data.medium.sel$accept

# calculate p.accept of simulated data
p.accept.mut = c()
p.accept.medium.sel = c()

# correct indexes
mut.site.mut = mut.site.mut - n.sites
mut.site.medium.sel = mut.site.medium.sel - n.sites

# calculate CN
CN.CM = CM.sim.data.mut$CN
CN = c()

for (i in (1:n.sites)) {
  CN[i] = CN.CM[which(mut.site.mut == i)][1]
}

for (i in (1:n.sites)) {
  accept.mut.i = accept.mut[which(mut.site.mut == i)]
  p.accept.mut.i = mean(accept.mut.i)
  p.accept.mut[i] = p.accept.mut.i

  accept.medium.sel.i = accept.medium.sel[which(mut.site.medium.sel == i)]
  p.accept.medium.sel.i = mean(accept.medium.sel.i)
  p.accept.medium.sel[i] = p.accept.medium.sel.i
}

## calculate corelations
r.p.accept.mut.CN = cor(p.accept.mut, CN)
r.p.accept.mut.score = cor(p.accept.mut, score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison CM" = c("log(paccept no-selection) vs CN", "log(paccept no-selection) vs score", "score vs CN"),
                    "R" = c(r.p.accept.mut.CN , r.p.accept.mut.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

print("<paccept no-selection>")
print(mean(p.accept.mut))

## calculate corelations
r.p.accept.medium.sel.CN = cor(log(p.accept.medium.sel), CN)
r.p.accept.medium.sel.score = cor(log(p.accept.medium.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison CM" = c("log(paccept selection) vs CN", "log(paccept selection) vs score", "score vs CN"),
                    "R" = c(r.p.accept.medium.sel.CN , r.p.accept.medium.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

print("<paccept selection>")
print(mean(p.accept.medium.sel))
```

```{r structure-profiles-CN-comparison}
d = d.CA
# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

## calculate correlations
r.exp.mean.z.RSD.CN = cor(d.exp$mean.z.RSD, CN)
r.mut.mean.z.RSD.CN = cor(d.mut$mean.z.RSD, CN)
r.medium.mean.z.RSD.CN = cor(d.medium$mean.z.RSD, CN)

cor.df = data.frame("comparison CA" = c("<z.RSD> experimental vs CN", "<z.RSD> no-selection vs CN", "<z.RSD> selection vs CN"),
                    "R" = c(r.exp.mean.z.RSD.CN , r.mut.mean.z.RSD.CN, r.medium.mean.z.RSD.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation structure (<z.RSD>) - sequence (CN)")
```

```{r structure-profiles-CN-comparison-CM}
d = d.CM
# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

## calculate correlations
r.exp.mean.z.RSD.CN = cor(d.exp$mean.z.RSD, CN)
r.mut.mean.z.RSD.CN = cor(d.mut$mean.z.RSD, CN)
r.medium.mean.z.RSD.CN = cor(d.medium$mean.z.RSD, CN)

cor.df = data.frame("comparison CM" = c("<z.RSD> experimental vs CN", "<z.RSD> no-selection vs CN", "<z.RSD> selection vs CN"),
                    "R" = c(r.exp.mean.z.RSD.CN , r.mut.mean.z.RSD.CN, r.medium.mean.z.RSD.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation structure (<z.RSD>) - sequence (CN)")
```

### Structure - normal modes

```{r set-file-names}
data.dir <- paste("OUT/out_subset_CA_ANM", sep = "")
R0 = R0.CA

# set input filenames
## profiles to compare
Pn.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.Pn.csv", sep = ""))
Pn.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
Pn.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
```

```{r prepare-site-data}
# prepare the data
Pn.exp = read.csv(Pn.exp.fname, header = T)
rownames(Pn.exp) = paste("mut", rownames(Pn.exp), sep = "")

Pn.mut = read.csv(Pn.mut.fname, header = T)
rownames(Pn.mut) = paste("mut", rownames(Pn.mut), sep = "")

Pn.medium = read.csv(Pn.medium.fname, header = T)
rownames(Pn.medium) = paste("medium", rownames(Pn.medium), sep = "")

mode = seq(ncol(Pn.medium))

# Prepare modeComparison.medium
tPn.medium = as.data.frame(t(as.matrix(Pn.medium)))
tPn.medium = cbind(data.frame("mode" = mode), tPn.medium)
modeComparison.medium = melt(tPn.medium, "mode", variable.name = "protein", value.name = "Pn")
modeComparison.medium = ddply(modeComparison.medium, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# Prepare modeComparison.mut
tPn.mut = as.data.frame(t(as.matrix(Pn.mut)))
tPn.mut = cbind(data.frame("mode" = mode), tPn.mut)
modeComparison.mut = melt(tPn.mut, "mode", variable.name = "protein", value.name = "Pn")
modeComparison.mut = ddply(modeComparison.mut, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# Prepare modeComparison.exp
tPn.exp = as.data.frame(t(as.matrix(Pn.exp)))
tPn.exp = cbind(data.frame("mode" = mode), tPn.exp)
modeComparison.exp = melt(tPn.exp, "mode", variable.name = "protein", value.name = "Pn")
modeComparison.exp = ddply(modeComparison.exp, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.medium <- na.omit(modeComparison.medium)
modeComparison.mut <- na.omit(modeComparison.mut)

CA.modeComparison.all.datasets = rbind(data.frame("dataset" = "exp", modeComparison.exp),
                                       data.frame("dataset" = "mut", modeComparison.mut),
                                       data.frame("dataset" = "medium", modeComparison.medium))
```

```{r median-vs-mode-energy}
dat = CA.modeComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

# calculate dataset of medians

dat = ddply(dat,c("dataset", "mode"), function(x) {
    Pn.median = median(x$Pn)
    Pn.median.ci = sort(x$Pn)[qbinom(c(.025, .975), length(x$Pn), 0.5)]
    Pn.median.min = Pn.median.ci[1]
    Pn.median.max = Pn.median.ci[2]
    data.frame(Pn.median.min, Pn.median.max, Pn.median)
})
```
```{r set-file-names}
data.dir <- paste("OUT/out_subset_CM_ANM", sep = "")
R0 = R0.CM

# set input filenames
## profiles to compare
Pn.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.Pn.csv", sep = ""))
Pn.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
Pn.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
```

```{r prepare-site-data}
# prepare the data
Pn.exp = read.csv(Pn.exp.fname, header = T)
rownames(Pn.exp) = paste("mut", rownames(Pn.exp), sep = "")

Pn.mut = read.csv(Pn.mut.fname, header = T)
rownames(Pn.mut) = paste("mut", rownames(Pn.mut), sep = "")

Pn.medium = read.csv(Pn.medium.fname, header = T)
rownames(Pn.medium) = paste("medium", rownames(Pn.medium), sep = "")

mode = seq(ncol(Pn.medium))

# Prepare modeComparison.medium
tPn.medium = as.data.frame(t(as.matrix(Pn.medium)))
tPn.medium = cbind(data.frame("mode" = mode), tPn.medium)
modeComparison.medium = melt(tPn.medium, "mode", variable.name = "protein", value.name = "Pn")
modeComparison.medium = ddply(modeComparison.medium, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# Prepare modeComparison.mut
tPn.mut = as.data.frame(t(as.matrix(Pn.mut)))
tPn.mut = cbind(data.frame("mode" = mode), tPn.mut)
modeComparison.mut = melt(tPn.mut, "mode", variable.name = "protein", value.name = "Pn")
modeComparison.mut = ddply(modeComparison.mut, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# Prepare modeComparison.exp
tPn.exp = as.data.frame(t(as.matrix(Pn.exp)))
tPn.exp = cbind(data.frame("mode" = mode), tPn.exp)
modeComparison.exp = melt(tPn.exp, "mode", variable.name = "protein", value.name = "Pn")
modeComparison.exp = ddply(modeComparison.exp, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.medium <- na.omit(modeComparison.medium)
modeComparison.mut <- na.omit(modeComparison.mut)

CM.modeComparison.all.datasets = rbind(data.frame("dataset" = "exp", modeComparison.exp),
                                       data.frame("dataset" = "mut", modeComparison.mut),
                                       data.frame("dataset" = "medium", modeComparison.medium))
```

```{r median-vs-mode-energy}
dat = CM.modeComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

# calculate dataset of medians

d = ddply(dat,c("dataset", "mode"), function(x) {
    Pn.median = median(x$Pn)
    Pn.median.ci = sort(x$Pn)[qbinom(c(.025, .975), length(x$Pn), 0.5)]
    Pn.median.min = Pn.median.ci[1]
    Pn.median.max = Pn.median.ci[2]
    data.frame(Pn.median.min, Pn.median.max, Pn.median)
})
```