### 2017-06-25 analysis-MSF-SD-RSD.Rmd
23/6/17

---
title: "Dynamical Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---
# Introduction
In this report we compare structural divergence profiles in cartesian coordinates (RMSD and MSD) with the Mean Square Fluctuation (MSF) of the sites.

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

# load functions
source(file.path(data.dir, "multiplot.R"))
source(file.path(data.dir, "my-functions.R")) 
```

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen","darkgreen", "red", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "medium", "selection", "mut+sel")
```

```{r set-run-variables}
# general parameters
tolerance = 1e-10

# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

## profiles to compare
### dri2
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
```

```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```
We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set-experimental-protein-names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r set-site-info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")

# add info to site.info
d = ddply(site.info, "site", .fun = function(x) data.frame("da" = min(x[, c(-1, -2)])))

# define shells of active.sites site neighborhood
d$shell = cut(d$da, breaks = c(-.1, 2.5, 7.5, 12.5, 17.5, 22.5, 27.5, 32.5, 37.5, 42.5, 47.5, 52.5, max(d$da)))
levels(d$shell) = seq(length(levels(d$shell))) - 1 # rename shell levels
site.info = merge(site.info, d, by = "site")

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
r.p.ref = ref.pdb$xyz[inds$xyz]
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno

pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

# add pdb.info to site.info
site.info = cbind(site.info, pdb.info)
```

```{r read-site-data}
# read dri2 data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)
```

```{r calculate-MSF-p.ref}
# calculate MSF of p.ref (Mean Square Fluctuation)
## calculate ENM
ENMK.p.ref = CalculateENMKeff(matrix(r.p.ref, nrow = 3), 
                              seq(1:n.sites), 
                              c(), 
                              R0, 
                              tolerance, 
                              K.analysis)
## get cov matrix
cov.p.ref = ENMK.p.ref$cov

## get the diagonal of the cov matrix
diag.p.ref = diag(cov.p.ref)

## calculate the factor to split the diagonal
factor = sort(rep(seq(1:n.sites), 3))

## split the diagonal
s.diag.p.ref = split(diag.p.ref, factor)

## create a matrix to save the data
MSF.p.ref = matrix(nrow = 1, ncol = n.sites)

## start a loop to calculate sums for each site
for (i in (1:n.sites)) {
  MSF.p.ref[, i] = sum(unlist(s.diag.p.ref[i]), use.names = F)
}

```

```{r prepare-site-data-dri2}
# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## mut
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.mut = cbind(site.info, tdri2.mut)
siteComparison.mut = melt(tdri2.mut, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## medium
tdri2.medium = as.data.frame(t(as.matrix(dri2.medium)))
tdri2.medium = cbind(site.info, tdri2.medium)
siteComparison.medium = melt(tdri2.medium, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets.dri2 = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                         data.frame("dataset" = "mut", siteComparison.mut),
                                         data.frame("dataset" = "medium", siteComparison.medium))

siteComparison.all.datasets.dri2 = ddply(siteComparison.all.datasets.dri2, c("dataset", "protein"), mutate,
          "RSD" = sqrt(dri2), 
           "SD" = dri2, 
        "z.RSD" = vscale(RSD),
         "z.SD" = vscale(SD))
```

```{r mean-profiles-comparisons}
# prepare dat
dat = siteComparison.all.datasets.dri2
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

# dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                mean.RSD = mean(x$RSD, na.rm = T)
                mean.z.SD = mean(x$z.SD, na.rm = T)
                mean.SD = mean(x$SD, na.rm = T)
                n.data = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(n.data)
                data.frame(mean.z.RSD,mean.RSD, mean.z.SD, mean.SD, n.data, se.z.RSD)
})

# extract means
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

# MSF.p.ref vs z.RSD and z.SD
MSF.p.ref = as.vector(MSF.p.ref)

# z.RSD
## correlations
r.exp = cor(d.exp$mean.RSD, MSF.p.ref)
r.medium = cor(d.mut$mean.RSD, MSF.p.ref)
r.mut = cor(d.medium$mean.RSD, MSF.p.ref)

cor.df = data.frame("MSF vs RSD" = c("exp", "selection", "no-selection"),
                             "R" = c(r.exp, r.medium, r.mut))
knitr::kable(cor.df, digits = 2)

## plots
par(pty = "s")

plot(y = d.exp$mean.RSD, x = MSF.p.ref, ylab = "RSD experimental")
lm = lm(d.exp$mean.RSD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.exp, digits = 2)), bty = "n")

plot(y = d.mut$mean.RSD, x = MSF.p.ref, ylab = "RSD no-selection")
lm = lm(d.mut$mean.RSD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.mut, digits = 2)), bty = "n")

plot(y = d.medium$mean.RSD, x = MSF.p.ref, ylab = "RSD selection")
lm = lm(d.medium$mean.RSD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.medium, digits = 2)), bty = "n")

# SD
## correlations
r.exp = cor(d.exp$mean.SD, MSF.p.ref)
r.medium = cor(d.mut$mean.SD, MSF.p.ref)
r.mut = cor(d.medium$mean.SD, MSF.p.ref)

cor.df = data.frame("MSF vs SD" = c("exp", "selection", "no-selection"),
                            "R" = c(r.exp, r.medium, r.mut))
knitr::kable(cor.df, digits = 2)

## plots
par(pty = "s")

plot(y = d.exp$mean.SD, x = MSF.p.ref, ylab = "SD experimental")
lm = lm(d.exp$mean.SD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.exp, digits = 2)), bty = "n")

plot(y = d.mut$mean.SD, x = MSF.p.ref, ylab = "SD no-selection")
lm = lm(d.mut$mean.SD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.mut, digits = 2)), bty = "n")

plot(y = d.medium$mean.SD, x = MSF.p.ref, ylab = "SD selection")
lm = lm(d.medium$mean.SD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.medium, digits = 2)), bty = "n")
```