### 2017-06-25 analysis-structure-MSD-RMSD-CN-WCN.Rmd

---
title: "Structural Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

# load functions
source("FUNCTIONS/ReadCA.R") 
source("FUNCTIONS/CalculateSideChainCM.R")
source(file.path(data.dir, "multiplot.R"))
source(file.path(data.dir, "my-functions.R"))
```

```{r set-file-names}
# set input filenames
cc.matrix = c()
s.cc.matrix = c()

# satart a loop for each family
for (f in (1:nrow(input))) { 
  print(f)
  family <- as.character(input$family)[f]
  type <- as.character(input$type)[f]
  p.ref <- as.character(input$p.ref)[f]
  enm <- as.character(input$enm)[f]
  n.mut.p <- input$n.mut.p[f]
  R0.CA = input$R0.CA[f]
  R0.CM = input$R0.CM[f]
  chain.p.ref <- as.character(input$chain.p.ref)[f]
  print(family)

  data.dir <- paste("OUT/out_subset_CA_ANM", sep = "")
  R0 = 10

  ## data
  reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
  protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))

  ## profiles to compare
  dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0,   "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
  dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
  dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}

# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom, ]
n.sites = length(ref.pdb.ca$resid)
AA = ref.pdb.ca$resid
resno = seq(1:n.sites)
pdb.info = data.frame(resno, AA)
names(pdb.info) = c("resno", "AA")
print(n.sites)

site.info = pdb.info

### CA ###

# calculate CN
R0 = 10
cmap.ca.R0.10 = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
CN = rowSums(cmap.ca.R0.10, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!

# calculate WCN
r.p.ref = matrix(ref.pdb$xyz[inds$xyz], nrow = 3)
dist.r.p.ref = dist.xyz(t(r.p.ref))
WCN = c()
for (i in (1:ncol(r.p.ref))) {
  WCN[i] = sum(1/dist.r.p.ref[i, -i]^2)
}

# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)
site = site.info$site
rownames(dri2.exp) = paste("exp", rownames(dri2.exp), sep = "")

## mut
dri2.mut = read.csv(dri2.mut.fname, header = T)
rownames(dri2.mut) = paste("mut", rownames(dri2.mut), sep = "")

## medium
dri2.medium = read.csv(dri2.medium.fname, header = T)
rownames(dri2.medium) = paste("medium", rownames(dri2.medium), sep = "")

# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## mut
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.mut = cbind(site.info, tdri2.mut)
siteComparison.mut = melt(tdri2.mut, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## medium
tdri2.medium = as.data.frame(t(as.matrix(dri2.medium)))
tdri2.medium = cbind(site.info, tdri2.medium)
siteComparison.medium = melt(tdri2.medium, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "medium", siteComparison.medium))

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
                     "z.SD" = vscale(dri2),
                      "RSD" = sqrt(dri2),
                    "z.RSD" = vscale(RSD))

## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

## dat with means and standard errors
d = ddply(dat,c("dataset", "resno"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                mean.z.SD = mean(x$z.SD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                data.frame(mean.z.RSD, mean.z.SD, ndata)
})

# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

# calculate correlations

cc = c()
cc = c(cc,cor(d.exp$mean.z.RSD, CN))
cc = c(cc,cor(d.exp$mean.z.RSD, WCN))
cc = c(cc,cor(d.exp$mean.z.RSD, 1/CN))
cc = c(cc,cor(d.exp$mean.z.RSD, 1/WCN))
cc = c(cc,cor(d.mut$mean.z.RSD, CN))
cc = c(cc,cor(d.mut$mean.z.RSD, WCN))
cc = c(cc,cor(d.mut$mean.z.RSD, 1/CN))
cc = c(cc,cor(d.mut$mean.z.RSD, 1/WCN))
cc = c(cc,cor(d.medium$mean.z.RSD, CN))
cc = c(cc,cor(d.medium$mean.z.RSD, WCN))
cc = c(cc,cor(d.medium$mean.z.RSD, 1/CN))
cc = c(cc,cor(d.medium$mean.z.RSD, 1/WCN))

cc = c(cc,cor(d.exp$mean.z.SD, CN))
cc = c(cc,cor(d.exp$mean.z.SD, WCN))
cc = c(cc,cor(d.exp$mean.z.SD, 1/CN))
cc = c(cc,cor(d.exp$mean.z.SD, 1/WCN))
cc = c(cc,cor(d.mut$mean.z.SD, CN))
cc = c(cc,cor(d.mut$mean.z.SD, WCN))
cc = c(cc,cor(d.mut$mean.z.SD, 1/CN))
cc = c(cc,cor(d.mut$mean.z.SD, 1/WCN))
cc = c(cc,cor(d.medium$mean.z.SD, CN))
cc = c(cc,cor(d.medium$mean.z.SD, WCN))
cc = c(cc,cor(d.medium$mean.z.SD, 1/CN))
cc = c(cc,cor(d.medium$mean.z.SD, 1/WCN))

cc.matrix = cbind(cc.matrix, cc)

cc = c()
cc = c(cc,cor(d.exp$mean.z.RSD, CN, method = "spearman"))
cc = c(cc,cor(d.exp$mean.z.RSD, WCN, method = "spearman"))
cc = c(cc,cor(d.exp$mean.z.RSD, 1/CN, method = "spearman"))
cc = c(cc,cor(d.exp$mean.z.RSD, 1/WCN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.RSD, CN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.RSD, WCN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.RSD, 1/CN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.RSD, 1/WCN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.RSD, CN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.RSD, WCN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.RSD, 1/CN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.RSD, 1/WCN, method = "spearman"))

cc = c(cc,cor(d.exp$mean.z.SD, CN, method = "spearman"))
cc = c(cc,cor(d.exp$mean.z.SD, WCN, method = "spearman"))
cc = c(cc,cor(d.exp$mean.z.SD, 1/CN, method = "spearman"))
cc = c(cc,cor(d.exp$mean.z.SD, 1/WCN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.SD, CN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.SD, WCN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.SD, 1/CN, method = "spearman"))
cc = c(cc,cor(d.mut$mean.z.SD, 1/WCN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.SD, CN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.SD, WCN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.SD, 1/CN, method = "spearman"))
cc = c(cc,cor(d.medium$mean.z.SD, 1/WCN, method = "spearman"))

s.cc.matrix = cbind(s.cc.matrix, cc)
}

colnames(cc.matrix) = input$family
colnames(s.cc.matrix) = input$family

names = c(                "exp.mean.z.RSD vs CN",
                        "exp.mean.z.RSD vs WCN",
                        "exp.mean.z.RSD vs 1/CN",
                        "exp.mean.z.RSD vs 1/WCN",
                        "no-sel.mean.z.RSD vs CN",
                        "no-sel.mean.z.RSD vs WCN",
                        "no-sel.mean.z.RSD vs 1/CN",
                        "no-sel.mean.z.RSD vs 1/WCN",
                        "sel.mean.z.RSD vs CN",
                        "sel.mean.z.RSD vs WCN",
                        "sel.mean.z.RSD vs 1/CN",
                        "sel.mean.z.RSD vs 1/WCN",
                        "exp.mean.z.SD vs CN",
                        "exp.mean.z.SD vs WCN",
                        "exp.mean.z.SD vs 1/CN",
                        "exp.mean.z.SD vs 1/WCN",
                        "no-sel.mean.z.SD vs CN",
                        "no-sel.mean.z.SD vs WCN",
                        "no-sel.mean.z.SD vs 1/CN",
                        "no-sel.mean.z.SD vs 1/WCN",
                        "sel.mean.z.SD vs CN",
                        "sel.mean.z.SD vs WCN",
                        "sel.mean.z.SD vs 1/CN",
                        "sel.mean.z.SD vs 1/WCN")
rownames(cc.matrix) = names
rownames(s.cc.matrix) = names

write.csv(cc.matrix, file="pearson.cc.SDvsPD.csv")
write.csv(s.cc.matrix, file="spearman.cc.SDvsPD.csv")

```

### All correlations

```{r correlations}
# build a data frame
cor = data.frame("CA pearson" = CA.cor.p,
                 "CM pearson" = CM.cor.p,
                 "CA spearman" = CA.cor.s,
                 "CM spearman" = CM.cor.s)
rownames(cor) = c("CN R0 = -0.1", "CN R0 = 2.5", "CN R0 = 7.5", "CN R0 = 12.5", "CN R0 = 17.5", "CN R0 = 22.5", "CN R0 = 27.5", "CN R0 = 32.5", "CN R0 = 37.5", "CN R0 = 42.5", "CN R0 = 47.5", "CN R0 = 52.5", "WCN")
knitr::kable(cor, digits = 2, caption = "Correlation structure (<z.RSD>) - sequence (CN/WCN)")
```